---
title: "HW 1-Categorical Data"
output: html_document
---

Agresti Problems

1.1
a. Response variable: attitude towards gun control 
   Explanatory variables: gender, mother's education 
b. Response variable: heart disease
   Explanatory variable: blood pressure, cholesterol level
c. Response variable: Vote for president
   Explanatory variables: annual income, race, religion
d. Response variable: Quality of life
   Explanatory variable: Marital status

1.2 

Which scale of measurement is most appropriate for the following variables -
nominal, or ordinal?
a. Political party affiliation-nominal
b. Highest degree obtained- ordinal 
c. Patient condition-ordinal
d. Hospital location-nominal
e. Favorite beverage-nominal 
f. How often feel depressed-ordinal


MGS1.1 

Multiple Choice Random Guessing Each of 20 multiple-choice questions on an exam has four possible answers but only one correct response. For each question, a student randomly selects one response as the answer.


a. Distribution of the student's number of correct answers on the exam. (Remember to specify the values of the parameters).

$Y \sim Binomial \big( 20, 0.25 \big)$

b. Generate a table of probabilities containing a list of each of the 21 possible outcomes (0 to 20 answers correct). If any probability is less than 0.0001, you can just list it as "< 0.0001". Do not use "E" notation.

```{r}
probvec<-seq(0, 21,by=1)
results<-round(dbinom(probvec, 20, 0.25),digits=2)
probtab<-data.frame(probvec, results)
print(probtab)

dbinom (0, 20, 0.25)
```

c. Name the assumptions that go with this distribution. Do you think that they are satisfied?
1. Each trial is independent
2. Trials have only two outcomes
3. Each trial has an equal probability of success.

I think they are satisfied if the answers are completely at random and the result of one question doesn't influence the result of another question. 

d. What is the probability that the student gets 9 or more correct responses?
$Pr (Y \geq 9)= Pr (Y=9) + Pr (Y=10) + Pr (Y=11)+....+Pr (Y=20)= 0.04092517$

```{r}
dbinom (9,20, 0.25) + dbinom (10,20, 0.25) + dbinom (11,20, 0.25) + dbinom (12,20, 0.25) + dbinom (13,20, 0.25) + dbinom (14,20, 0.25) + dbinom (15,20, 0.25) + dbinom (16,20, 0.25) + dbinom (17,20, 0.25) + dbinom (18,20, 0.25) + dbinom (19,20, 0.25) + dbinom (20,20, 0.25)

probvec<-seq(9,21, by=1)
prob<-dbinom(probvec, 20, 0.25)
probsum<-sum(dbinom(probvec, 20, 0.25))
probsum
```


e. What is the probability that the student gets 2 or fewer questions correct?
$Pr (Y \leq 2)= Pr (Y=0) + Pr (Y=1) + Pr (Y=1)= (2,20,0.25)$

```{r}
pbinom (2,20,0.25)
```


f. What is the expected number of correct answers?
5 correct answers

$\mu (Y) = N\pi= 20*0.25=5$

MGS1.2 Coin flipping 

N= 9  
 ???????.
a. Assuming $ \pi$=????0.5, specify the probabilities for the possible values for $Y$????????, and find the distribution's mean and standard deviation.

```{r}
propvec<-seq (0,9, by=1)
prob<-round(dbinom(propvec, 9, 0.5), digits=2)
results<-data.frame(propvec, prob)
results


```

$\mu (Y) = N\pi= 9*0.5=4.5$

$\sigma (Y)= \sqrt N\pi (1-\pi)= \sqrt 9*.5 * 0.5= 1.5$


b. Find the binomial probabilities for $Y$ when $\pi$ = 0.6 and 0.3

```{r}
propvec<-seq(0,9, by=1)
prob<-round(dbinom(propvec,9,0.6), digits=2)
results<-data.frame (propvec, prob)
results
```

```{r}
propvec<-seq(0,9, by=1)
prob<-round(dbinom(propvec,9,0.3), digits=2)
results<-data.frame (propvec, prob)
results
```



c. Suppose that you have a coin that is "not fair" - meaning that the probability of a head is different from 0.5. You observe $y$????????=7 and do not know $\pi$????????. Use R to calculate and graph the likelihood function. What is the MLE of $\pi$????????? Show a graph of the likelihood function when
y????????=7

The MLE of $\pi$ when there were 7 heads in 9 trials is 0.780
```{r}
pi_vals <- seq(0,1,by=0.005)              
lhood <- dbinom(7,9,prob=pi_vals)        
plot(pi_vals,lhood, type="l", col="red",  # plot first likelihood
     main="Likelihoods with 9 trials",
     xlab = "pi", ylab="Likelihood")
lhoodtab<-data.frame(pi_vals,lhood)
lhoodtab2<-subset(lhoodtab, lhood > 0.3)
print(lhoodtab2)
MLE<-round(7/9, digits=2)
```


MGS1.3 Sunblock Lotion Each of 12 volunteers used a new skin product that is purported to improve dry skin. At the end of the study 10 out of 12 patients showed improvement. Let $\pi$????????= the true probability of improvement. Our null hypothesis is $H_0???: $????????=0.5. Use a 95% significance level for all tests.

a. Using the binomial test, do a hypothesis test at the 5% level (do all steps) and find the mid p-value for H ????????????????:???????????0.5 . What is your conclusion? (Write a statistical conclusion and a "plain English" conclusion.)
Research question: Does using the sunblock lotion help improve dry skin? 
$H_0$: $\pi=0.5$
$H_A$: $\pi \neq 0.5$
If we assume the $H_0$ is true, the probability of seeing 10/12 patients with improvement would be:0.0161. With this in mind, the probability of seeing a result as extreme or more extreme when $\pi$=0.5, with mid-p corection  is = Prob(Y=0) + Prob(Y=1) + 0.5(Prob(Y=2)) + 0.5(Prob(Y=10) + Prob(Y=11) + Prob(Y=12)= 0.022

Conclusion: At the 95% level, we reject the null hypothesis. The evidence suggests that the probability of skin improvement using the sunblock lotion is not equal to 0.5.

```{r}
numsucc<-seq(0,12, by=1)
probsucc<-dbinom(numsucc,12,0.5)
plot (numsucc,probsucc, type="l")
probtab<-round(data.frame(numsucc,probsucc),4)
probtab
P_val= sum(dbinom(0,12,0.5) + dbinom(1,12,0.5) +0.5*dbinom(2,12,0.5) +0.5*dbinom(10,12,0.5) + dbinom(11,12,0.5) + dbinom(12,12,0.5))
P_val

#dbinom(10,12,0.5)
#1-pbinom (10, 12, 0.5)
```
b. Using the binomial test, do a hypothesis test at the 5% level (do all steps) and find the mid p-value for $H_A:\pi$????????????????:????????>0.5 . What is your conclusion? (Write a statistical conclusion and a "plain English" conclusion.)
Research question: is the probability of improving dry skin when using the lotion higher than 5?
$H_0:\pi\leq 5$
$H_A: \pi>5$
The probability of seeing 10 or more people improving assuming $H_0$ is true is= 0.5 Prob(Y=10) + Prob(Y=11) +  Prob(Y=12)= 0.011.
Conclusion: at the 95% level, we reject the null hypothesis. The evidence suggets that the probability of improving dry skin using the lotion is greater than 0.5.

```{r}
p_value=0.5*dbinom(10, 12, 0.5) + dbinom(11, 12, 0.5) + dbinom (12, 12, 0.5)
p_value
```


c. What is the Agresti-Coull 95% confidence interval for $\pi$

```{r}
y=10
n=12
p_prime=(y+2)/(n+4)
z=abs(qnorm(0.05/2))
stat=(p_prime * (1-p_prime))/(n+4)
Al=p_prime - z*(sqrt(stat))
Al
Au= p_prime + z*(sqrt(stat))
Au
```
The Agresti-Coull 95% CI is (0.538,0.962)

????
d. What is the Score (inversion) 99% confidence interval for ?????????
```{r}
n=12
pvec<-seq(0,1,by=0.01)
p=10/12
z_stat<-((p-pvec)/(sqrt(pvec*(1-pvec))/n))
threshold= abs(qnorm(0.01/2))
prob_tab<-data.frame (pvec, z_stat)
prob_tab_2<-prob_tab[abs(z_stat)<=abs(threshold),]
prob_tab_2
```
The Score 99% CI is (0.74,0.89)


MGS1.4 Flu Vaccinations On the first day of class, we did a poll and found that 16 out of 44 students had received their flu vaccination for the 2018-19 flu season. Among all US residents of the 18-49 age group, 33.6% have been vaccinated. Assume that our class is a random sample of Cornell students.

a. Do a hypothesis test, using the Score statistic, to test of the hypothesis that Cornell students are vaccinated at a rate of 33.6%, versus a two-sided alternative. Be sure in include a "plain English" conclusion. Perform the test at the $\alpha$????????=5% level.
Research question: is the proportion of students who received the flu vaccine at Cornell similar to the national average?
$H_0: \pi=\pi_0=0.336$
$H_A: \pi\neq 0.336$
Test statistic= 0.388

```{r}
n=44
p=16/44
p_0=0.336
#z=((p-p_0)/(sqrt((p_0*(1-p_0)/n))
za= p-p_0
zb= (p_0*(1-p_0))/n
z=za/sqrt(zb)
1-pnorm(z)
```

b. Calculate the Agresti-Coull confidence interval for the true vaccination proportion, based on our class sample. Does this interval support the conclusion in part (a)?
```{r}
n=44
y=16
pp=(y+2)/(n+4)
z=abs(qnorm(0.05/2))
stat=sqrt((pp*(1-pp))/(n+4))
Al= pp - z*stat
Au=pp + z*stat
```
The Agresti-Coull 95% CI for the true vaccination proportion is (0.238,0.512). This interval includes $\pi_0$, so we fail to reject the null hypothesis. This result supports the conclusion in part a.

MGS 1.5 Coverage Probabilities Your task is to use simulation to estimate the coverage probabilities of Wald and Agresti-Coull confidence intervals in four different situations:
1)$\pi$????????=0.1,????????=15
2)$\pi$????????=0.1,????????=75
3)$\pi$????????=0.55,????????=15
4 $\pi$????????=0.55,????????=75
You should use 10,000 simulations of each of the above situations, to find the coverage probabilities. Please include your R code. Write your answers in the table below.
1)
```{r}
n=15
pi=0.1
sim<-rbinom (10000,n, pi)
p<-(sim/n)
pp<-(sim+2)/(n+4)
z=abs(qnorm(0.05/2))
stat=sqrt((pp*(1-pp))/(n+4))
Al=abs(pp - (z*stat))
Au=pp + (z*stat)
result<- p >= Al & p <= Au
count_true<-length(which(result=="TRUE"))
count_true/10000
```

```{r}
n=15
pi=0.1
sim<-rbinom (10000,n, pi)
p<-(sim/n)
z=abs(qnorm(0.05/2))
stat=sqrt((p*(1-p))/n)
Wl=abs(p - (z*stat))
Wu=p + (z*stat)
result<- p >= Wl & p <= Wu
count_true<-length(which(result=="TRUE"))
count_true/10000
```



